---
title: "Hybrid_Teaching_Regression"
output:
  html_document:
    df_print: paged
---
#Importing required libraries
```{r}
library(readxl)
library(raincloudplots)
library(ggrain)
library(table1)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(corrplot)
library(stats)
library(rpart)
library(rpart.plot)
library(caret)
library(e1071)
library(writexl)
```


#importing the data
```{r}
data <- read_excel("C:\\Users\\Angat\\OneDrive\\Desktop\\STM5IPL_HYBRID_LEARNING\\Deidentified Data Detailed Complie-03-02-2023 (no qual).xlsx", sheet = "Matched Data Compile 12.11.22")
data<-data[-89,]
data <- subset(data, select = -c(...2,...66))
data
```

#selecting relevant columns for analysis
```{r}
selected_columns <-subset(data, select =c("Mode", "Region origin", "Campus", "ATAR", "Course Weighted Average","Age", "Gender", "IMI Total","IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness","SUS", "AT2 article summary Wk 5?","AT 3: Critical appraisal essay Wk9", "IM_Percentage", "EM_Percentage","Motivation_Type", "Like_Hybrid","TOTAL","AT3_Performance","AT3_Median_Performance","TOTAL_Performance"))

# Display the names of the selected columns
names(selected_columns)

```

#Data Cleaning
```{r}
#Check the ATAR with 0 values
data[data$ATAR==0,]
```


```{r}
# Replace 0 values in ATAR column with "NA"
selected_columns$ATAR <- ifelse(selected_columns$ATAR == 0,NA, selected_columns$ATAR)

# Replace "-" with NA in specific columns using mutate_at
selected_columns <- selected_columns %>%
  mutate_at(
    vars( `AT2 article summary Wk 5?`, `AT 3: Critical appraisal essay Wk9`),
    ~ifelse(. %in% c("-", " ","0"), NA, .)
  )



```





#Data Exploration
```{r}
my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
  
}

# Generate Table 1
table1(
  ~ Age + ATAR + `Course Weighted Average` +`IMI Total`+`SUS` + `AT2 article summary Wk 5?` + `AT 3: Critical appraisal essay Wk9` +`IM_Percentage` + `EM_Percentage`+`Like_Hybrid` | Mode,
  data=selected_columns,
  render.continuous = my.render.cont,
  caption="Table 1 : Descriptive statistics exploration for all variables wrt Study Mode"
)


```

```{r}
#Stat summary for Mode

table1(~Age + ATAR + `Course Weighted Average`+`SUS`+`AT2 article summary Wk 5?`+`AT 3: Critical appraisal essay Wk9`+`IM_Percentage`+`EM_Percentage`|Mode,data=selected_columns,caption="Table 1 : Descriptive statistics for Mode",render.continuous = my.render.cont)
```


```{r}
#Stat summary for Mode and Motivation Type

table1(~Age + ATAR + `Course Weighted Average`+`SUS`+`AT2 article summary Wk 5?`+`AT 3: Critical appraisal essay Wk9`+`IM_Percentage`+`EM_Percentage`|Mode*Motivation_Type,data=selected_columns,caption="Table 2 : Descriptive statistics for Mode and Motivation type",render.continuous = my.render.cont)
```



```{r}
#Stat summary for Mode and Gender

table1(~Age + ATAR + `Course Weighted Average`+`SUS`+`AT2 article summary Wk 5?`+`AT 3: Critical appraisal essay Wk9`+`IM_Percentage`+`EM_Percentage`|Mode*Gender,data=selected_columns,caption="Table 3 : Descriptive statistics for Mode and Gender",render.continuous = my.render.cont)
```


```{r}
#Stat summary for Motivation type and Gender

table1(~Age + ATAR + `Course Weighted Average`+`SUS`+`AT2 article summary Wk 5?`+`AT 3: Critical appraisal essay Wk9`+`IM_Percentage`+`EM_Percentage`|Motivation_Type*Gender,data=selected_columns,caption="Table 4 : Descriptive statistics for Motivation Type and Gender")
```


#Checking for missing values after imputation of null values
```{r}
relevant_cols <- c("Mode", "Region origin", "Campus", "ATAR", "Course Weighted Average","Age", "Gender", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI Total","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness","SUS", "AT2 article summary Wk 5?","AT 3: Critical appraisal essay Wk9", "IM_Percentage", "EM_Percentage","Motivation_Type", "Like_Hybrid","TOTAL","AT3_Performance","AT3_Median_Performance","TOTAL_Performance")
complete_data <- selected_columns[complete.cases(selected_columns[, relevant_cols]), ]
colSums(is.na(complete_data))
```

```{r}
# Create a count plot for Motivation_Type
ggplot(complete_data, aes(y = Motivation_Type)) +
  geom_bar(fill = "#3676e8", color = "black") +
  labs(y = "Motivation Type", x = "Count") +
  ggtitle("Count of Motivation Types")

```


```{r}
# Create a count plot for Motivation_Type
ggplot(complete_data, aes(y = Like_Hybrid)) +
  geom_bar(fill = "#ff6767", color = "black") +
  labs(y = "Like Hybrid", x = "Count") +
  ggtitle("Count of Students Liking Hybrid")
```


```{r}
like_hybrid_yes<-dplyr::filter(complete_data,Like_Hybrid=="YES")
like_hybrid_yes<-dplyr::select(like_hybrid_yes,Motivation_Type)
like_hybrid_yes<-dplyr::count(like_hybrid_yes,Motivation_Type)
intrinsic_count <- like_hybrid_yes$n[like_hybrid_yes$Motivation_Type == "INTRINSIC"]
extrinsic_count <- like_hybrid_yes$n[like_hybrid_yes$Motivation_Type == "EXTRINSIC"]
ratio <- intrinsic_count / extrinsic_count
print(ratio)

# Create the first plot
plot1 <- ggplot(complete_data, aes(y = Motivation_Type,fill=Motivation_Type)) +
  geom_bar(position = "dodge", color = "black") +
  labs(y = "Motivation Type", x = "Count") +scale_fill_manual(values = c("INTRINSIC" = "#a8dbad", "EXTRINSIC" = "#3498db"))+
  ggtitle("Count of Motivation Types")

# Create the second plot
plot2 <- ggplot(complete_data, aes(y = Motivation_Type, fill = Like_Hybrid)) +
  geom_bar(position = "dodge", color = "black") +
  labs(y = "Motivation Type", x = "Count") +
  ggtitle("Ratio of Intrinsic to Extrinsic motivated students") +
  scale_fill_manual(values = c("YES" = "#3498db", "NO" = "#e30022"))

# Arrange the plots in a single frame
grid.arrange(plot1, plot2, nrow = 2)

```

```{r}
plot.1<-ggplot(complete_data, aes(y = Mode, x = `AT 3: Critical appraisal essay Wk9`, fill = Mode)) +
  geom_boxplot() +
  labs(y = "Mode", x = "AT 3: Critical appraisal essay Wk9") +
  ggtitle("AT3 Grade distribution for different Modes") +
  theme_minimal() +
  scale_fill_manual(values = c("Hybrid/Class.com" = "#3498db", "Online/Zoom" = "#a8dbad"))  # Assuming these are the


plot.2<-ggplot(complete_data, aes(y = Motivation_Type, x = `AT 3: Critical appraisal essay Wk9`, fill = Motivation_Type)) +
  geom_boxplot() +
  labs(y = "Motivation Type", x = "AT 3: Critical appraisal essay Wk9") +
  ggtitle("AT3 Grade distribution for different Motivation Types") +
  theme_minimal() +
  scale_fill_manual(values = c("INTRINSIC" = "#a8dbad", "EXTRINSIC" = "#3498db"))  # Assign colors as per your

# Arrange the plots in a single frame
plt.3<-grid.arrange(plot.1, plot.2, nrow = 2)
```

```{r}
# Assuming 'Mode' is a factor variable
plt.4<-ggplot(complete_data, aes(x = Gender, y = `AT 3: Critical appraisal essay Wk9`, fill = Gender)) +
  geom_boxplot() +
  labs(x = "Gender", y = "AT 3: Critical appraisal essay Wk9") +
  ggtitle("Distribution of 'AT 3: Critical appraisal essay Wk9' by Gender") +
  theme_minimal()+ggtitle("AT3 distribution for different Gender")+
  scale_fill_manual(values = c("Female" = "pink", "Male" = "#add8e6"))

grid.arrange(plt.3,plt.4,ncol=2)
```


```{r}
# Box plot for Gender
p1 <- ggplot(complete_data, aes(x = Gender, y = `AT 3: Critical appraisal essay Wk9`, fill = Gender)) +
  geom_boxplot() +
  labs(y = "Gender", x = "AT 3: Critical appraisal essay Wk9") +
  ggtitle("Distribution of AT 3 by Gender") +
  scale_fill_manual(values = c("Female" = "pink", "Male" = "#add8e6"))+
  theme_minimal()

# Box plot for Like_Hybrid
p2 <- ggplot(complete_data, aes(x = as.factor(Like_Hybrid), y = `AT 3: Critical appraisal essay Wk9`, fill = Like_Hybrid)) +
  geom_boxplot() +
  labs(y = "Like_Hybrid", x = "AT 3: Critical appraisal essay Wk9") +
  ggtitle("Distribution of 'AT 3 by Like_Hybrid") +
  scale_fill_manual(values = c("NO" = "#e30022", "YES" = "#a8dbad"))+
  theme_minimal()

# Arrange the two plots in a 2x1 grid
plot.3<-grid.arrange(p1, p2, ncol = 2)
grid.arrange(plt.3,plot.3,nrow=2)

```

```{r}
# List of numerical variable names with double quotes
numerical_vars <- c("ATAR", "Course Weighted Average", "Age", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness","SUS", "AT2 article summary Wk 5?", "AT 3: Critical appraisal essay Wk9","TOTAL","IM_Percentage", "EM_Percentage")

# Reshape the data into a long format
data_long <- complete_data %>%
  pivot_longer(cols = all_of(numerical_vars))

# Create a grid of histograms
hist_grid <- ggplot(data_long, aes(x = value)) +
  geom_boxplot(fill = "#a8dbad", color = "black") +
  facet_wrap(~ name, scales = "free") +
  labs(x = "Value", y = "Frequency") +
  theme_minimal()+ggtitle("Outlier Analysis for Variables")

# Print the histogram grid
print(hist_grid)
```


#For Hybrid
```{r}
# Load the required packages
library(ggplot2)
library(tidyr)

# Assuming your data is loaded into a dataframe called 'data'
# Replace 'data' with your actual dataframe name

# List of numerical variable names with double quotes
numerical_vars <- c("ATAR", "Course Weighted Average", "Age", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness","SUS", "AT2 article summary Wk 5?", "AT 3: Critical appraisal essay Wk9","TOTAL","IM_Percentage", "EM_Percentage")

# Reshape the data into a long format
data_long <- complete_data %>%filter(Mode=="Hybrid/Class.com")%>%pivot_longer(cols = all_of(numerical_vars))

# Create a grid of histograms
hist_grid <- ggplot(data_long, aes(x = value)) +
  geom_boxplot(fill = "#4169e1", color = "black") +
  facet_wrap(~ name, scales = "free") +
  labs(x = "Value", y = "Frequency") +
  theme_minimal()+ggtitle("Outlier Analysis for Variables For Hybrid Mode")

# Print the histogram grid
print(hist_grid)
```

#Outlier analysis for Online Mode
```{r}
# Load the required packages
library(ggplot2)
library(tidyr)

# Assuming your data is loaded into a dataframe called 'data'
# Replace 'data' with your actual dataframe name

# List of numerical variable names with double quotes
numerical_vars <- c("ATAR", "Course Weighted Average", "Age", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness","SUS", "AT2 article summary Wk 5?", "AT 3: Critical appraisal essay Wk9","TOTAL","IM_Percentage", "EM_Percentage")

# Reshape the data into a long format
data_long <- complete_data %>%filter(Mode=="Online/Zoom")%>%pivot_longer(cols = all_of(numerical_vars))

# Create a grid of histograms
hist_grid <- ggplot(data_long, aes(x = value)) +
  geom_boxplot(fill = "#e8f48c", color = "black") +
  facet_wrap(~ name, scales = "free") +
  labs(x = "Value", y = "Frequency") +
  theme_minimal()+ggtitle("Outlier Analysis for Variables For Online Mode")

# Print the histogram grid
print(hist_grid)
```

#Outlier analysis for INTRINSIC motivation type
```{r}
# Load the required packages
library(ggplot2)
library(tidyr)

# Assuming your data is loaded into a dataframe called 'data'
# Replace 'data' with your actual dataframe name

# List of numerical variable names with double quotes
numerical_vars <- c("ATAR", "Course Weighted Average", "Age", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness","SUS", "AT2 article summary Wk 5?", "AT 3: Critical appraisal essay Wk9","TOTAL","IM_Percentage", "EM_Percentage")

# Reshape the data into a long format
data_long <- complete_data %>%filter(Motivation_Type=="INTRINSIC")%>%pivot_longer(cols = all_of(numerical_vars))

# Create a grid of histograms
hist_grid <- ggplot(data_long, aes(x = value)) +
  geom_boxplot(fill = "#6f2da8", color = "black") +
  facet_wrap(~ name, scales = "free") +
  labs(x = "Value", y = "Frequency") +
  theme_minimal()+ggtitle("Outlier Analysis for Variables Intrinsic Motivation")

# Print the histogram grid
print(hist_grid)
```

```{r}
library(psych)
describe(complete_data[,c("ATAR", "Course Weighted Average","Age","IMI Total","AT2 article summary Wk 5?","AT 3: Critical appraisal essay Wk9","IM_Percentage", "EM_Percentage","TOTAL")])
```

```{r}
#storing the cleaned dataset
write_xlsx(complete_data, path = "C:\\Users\\Angat\\Downloads\\hybrid_teaching_cleaned.xlsx")
```



#Statistical testing
```{r}
#Statistical test to check pvalues for different groups for AT3
atar_hybrid <- complete_data$Age[complete_data$Mode == "Hybrid/Class.com"]
atar_online <- complete_data$Age[complete_data$Mode == "Online/Zoom"]

# Perform Mann-Whitney U test
result <- wilcox.test(atar_hybrid, atar_online)

# Print the result
print(result$p.value)

#Like Mode we can use different col names and their values to test against AT3
```

```{r}
perform_stat_test <- function(data, variable, grouping_variable) {
  # Shapiro-Wilk Test for Normality
  shapiro_test_result <- shapiro.test(data[[variable]])
  
  if (shapiro_test_result$p.value < 0.05) {
    cat("For ",grouping_variable,"\n")
    cat("The variable", variable, "is not normally distributed.\n p-value:", shapiro_test_result$p.value)
    
    # Mann-Whitney U Test
    mwu_test_result <- wilcox.test(data[[variable]] ~ data[[grouping_variable]])
    
    # Print the Mann-Whitney U test results
    print(mwu_test_result)
    
    # Provide insights based on p-value
    if (mwu_test_result$p.value < 0.05) {
      cat("Based on the Mann-Whitney U test, there is a statistically significant difference in", variable, "between groups for .\n")
    } else {
      cat("Based on the Mann-Whitney U test, there is no statistically significant difference in", variable, "between groups.\n")
    }
  } else {
    cat("The variable", variable, "is normally distributed.\n p-value:", shapiro_test_result$p.value)
    
    # Independent Samples t-test
    t_test_result <- t.test(data[[variable]] ~ data[[grouping_variable]])
    
    # Print the t-test results
    print(t_test_result)
    
    # Provide insights based on p-value
    if (t_test_result$p.value < 0.05) {
      cat("Based on the Independent Samples t-test, there is a statistically significant difference in", variable, "between groups.\n")
    } else {
      cat("Based on the Independent Samples t-test, there is no statistically significant difference in", variable, "between groups.\n")
    }
  }
}

#Stat test between Mode and AT3
perform_stat_test(complete_data, "AT 3: Critical appraisal essay Wk9", "Mode")

perform_stat_test(complete_data, "AT 3: Critical appraisal essay Wk9", "Motivation_Type")

perform_stat_test(complete_data, "AT 3: Critical appraisal essay Wk9","Like_Hybrid")

perform_stat_test(complete_data, "AT 3: Critical appraisal essay Wk9","Gender")

# Filter the data for Intrinsic motivation type
intrinsic_data <- complete_data %>%
  filter(Motivation_Type == "INTRINSIC")

# Perform your statistical test on the filtered data
perform_stat_test(intrinsic_data, "AT 3: Critical appraisal essay Wk9", "Mode")

# Filter the data for Extrinsic motivation type
extrinsic_data <- complete_data %>%
  filter(Motivation_Type == "EXTRINSIC")

# Perform your statistical test on the filtered data
perform_stat_test(extrinsic_data, "AT 3: Critical appraisal essay Wk9", "Mode")

# Filter the data for Like Hybrid
love_hybrid <- complete_data %>%
  filter(Like_Hybrid == "YES")

# Perform your statistical test on the filtered data
perform_stat_test(love_hybrid, "AT 3: Critical appraisal essay Wk9", "Mode")



# Filter the data for dont Like Hybrid
dont_love_hybrid <- complete_data %>%
  filter(Like_Hybrid == "NO")

# Perform your statistical test on the filtered data
perform_stat_test(dont_love_hybrid, "AT 3: Critical appraisal essay Wk9", "Mode")


# Filter the data for dont Like Hybrid
hybrid_learning <- complete_data %>%
  filter(Mode == "Hybrid/Class.com")
#Stat test between Gender and AT3
perform_stat_test(hybrid_learning, "AT 3: Critical appraisal essay Wk9", "Gender")
```


```{r}
table <- table(complete_data$Mode,complete_data$Motivation_Type)

# Chi-square test
test_result <- fisher.test(table)

# Print the result
print(test_result)
```

```{r}
table <- table(complete_data$Mode,complete_data$Like_Hybrid)

# Chi-square test
test_result <- fisher.test(table)

# Print the result
print(test_result)
```

```{r}
table <- table(complete_data$Like_Hybrid,complete_data$Motivation_Type)

# Chi-square test
test_result <- fisher.test(table)

# Print the result
print(test_result)
```

```{r}
table <- table(complete_data$Gender,complete_data$Motivation_Type)

# Chi-square test
test_result <- fisher.test(table)

# Print the result
print(test_result)
```

```{r}
write.csv(complete_data, file ="C:\\Users\\Angat\\OneDrive\\Desktop\\STM5IPL_HYBRID_LEARNING\\hybrid_dataset.csv", row.names = FALSE)
```



#Transforming nominal categorical variables
```{r}
complete_data$Gender <- factor(complete_data$Gender, labels = c(0, 1))
complete_data$Campus_transformed <- as.numeric(factor(ifelse(complete_data$Campus == "Bundoora", 0, 1) ) )
complete_data$Age_Group <- ifelse(complete_data$Age <= 19,1,0)
```


#Transforming binary categorical variables
```{r}
regression_data<-complete_data
regression_data$Mode<- as.numeric(factor(regression_data$Mode, labels = c(0, 1)))
regression_data$Gender <- as.numeric(factor(regression_data$Gender, labels = c(0, 1)))
regression_data$Motivation_Type<- as.numeric(factor(regression_data$Motivation_Type, labels = c(0, 1)))
regression_data$Like_Hybrid<- as.numeric(factor(regression_data$Like_Hybrid, labels = c(0, 1)))

```


```{r}
# List of numerical variable names with double quotes
numerical_vars <- c("ATAR", "Course Weighted Average", "Age", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness", "AT2 article summary Wk 5?", "AT 3: Critical appraisal essay Wk9","TOTAL","IM_Percentage", "EM_Percentage")

# Reshape the data into a long format
data_long <- regression_data %>%
  pivot_longer(cols = all_of(numerical_vars))

# Create a grid of histograms
hist_grid <- ggplot(data_long, aes(x = value)) +
  geom_histogram(bins = 20, fill = "#3498db", color = "black") +
  facet_wrap(~ name, scales = "free") +
  labs(x = "Value", y = "Frequency") +
  ggtitle("Distribution of Variables")+
  theme_minimal()

# Print the histogram grid
print(hist_grid)
```


```{r}
# Assuming your data is loaded into a dataframe called 'regression_data'
# Replace 'regression_data' with your actual dataframe name

# Add a unique identifier to the original data
regression_data <- regression_data %>% mutate(id = row_number())

numerical_vars <- c("ATAR", "Course Weighted Average", "Age", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness", "AT2 article summary Wk 5?","IM_Percentage", "EM_Percentage")

# Reshape the data into a long format and add a unique identifier
data_long <- regression_data %>%
  pivot_longer(cols = all_of(numerical_vars), names_to = "Variable", values_to = "Value") %>%
  mutate(id = rep(regression_data$id, length(numerical_vars)))

# Join with the original data to bring in the "AT 3: Critical appraisal essay Wk9" values
data_long <- left_join(data_long, dplyr::select(regression_data, id, `AT 3: Critical appraisal essay Wk9`), by = "id")

target<-data_long$`AT 3: Critical appraisal essay Wk9.x`

# Create a grid of scatter plots with linear lines
scatter_grid <- ggplot(data_long, aes(x = target, y = Value)) +
  geom_point(aes(color = Variable), alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, aes(color = Variable)) +
  facet_wrap(~ Variable, scales = "free") +
  labs(x = "AT 3: Critical appraisal essay Wk9", y = "Value") +
  ggtitle("Correlation Between Independent and AT3 variable",subtitle = "Assesing linear relationship to confirm Regression assumption")
  theme_minimal()

# Print the scatter plot grid
print(scatter_grid)
```



```{r}
# List of numerical variable names with double quotes
numerical_vars <- c("ATAR", "Course Weighted Average", "Age", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness","SUS", "AT2 article summary Wk 5?","TOTAL","IM_Percentage", "EM_Percentage")


regression_data$`AT 3: Critical appraisal essay Wk9` <- log(regression_data$`AT 3: Critical appraisal essay Wk9`)
```


```{r}
# Load the required packages
# Assuming your data is loaded into a dataframe called 'regression_data'
# Replace 'regression_data' with your actual dataframe name

# Add a unique identifier to the original data
regression_data <- regression_data %>% mutate(id = row_number())

numerical_vars <- c("ATAR", "Course Weighted Average", "Age", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness", "AT2 article summary Wk 5?","IM_Percentage", "EM_Percentage")

# Reshape the data into a long format and add a unique identifier
data_long <- regression_data %>%
  pivot_longer(cols = all_of(numerical_vars), names_to = "Variable", values_to = "Value") %>%
  mutate(id = rep(regression_data$id, length(numerical_vars)))

# Join with the original data to bring in the "AT 3: Critical appraisal essay Wk9" values
data_long <- left_join(data_long, dplyr::select(regression_data, id, `AT 3: Critical appraisal essay Wk9`), by = "id")

target<-data_long$`AT 3: Critical appraisal essay Wk9.x`

# Create a grid of scatter plots with linear lines
scatter_grid <- ggplot(data_long, aes(x = target, y = Value)) +
  geom_point(aes(color = Variable), alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, aes(color = Variable)) +
  facet_wrap(~ Variable, scales = "free") +
  labs(x = "AT 3: Critical appraisal essay Wk9", y = "Value") +
  ggtitle("Correlation Between Independent and AT3 variable",subtitle = "Assesing linear relationship to confirm Regression assumption After log transformation")
  theme_minimal()

# Print the scatter plot grid
print(scatter_grid)


```


```{r}
#Removing unwanted variables for linear regression modelling to avoid multicollinearity
lin_regression_data<-dplyr::select(regression_data,-c(Campus,`Region origin`,AT3_Performance,AT3_Median_Performance,TOTAL,TOTAL_Performance,SUS,`IMI - Interest/Enjoyment`,`IMI - Perceived Competence`,`IMI - Effort/Importance`,`IMI - Perceived Choice`,`IMI - Value/Usefulness`,`IMI - Relatedness`,`IMI Total`,EM_Percentage,Like_Hybrid,id,`Course Weighted Average`))

numerical_vars <- c("ATAR", "Course Weighted Average", "Age_Group", "IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness", "AT2 article summary Wk 5?","IM_Percentage","EM_Percentage","TOTAL","Gender","Motivation_Type","Campus_transformed","AT 3: Critical appraisal essay Wk9")

# Calculate the correlation matrix
correlation_matrix <- cor(regression_data[,numerical_vars])


# Create a correlation matrix plot using corrplot with labels
corrplot(correlation_matrix, method = "color", type = "upper", tl.col = "black", tl.srt = 45,
         tl.cex = 0.7, addCoef.col = "black", number.cex = 0.5)
```

#Correlation plot after removing independent variables that were highly correlated among themsleves
```{r}
numerical_vars <- c("Mode","ATAR",
    "Age_Group","Gender","AT2 article summary Wk 5?","IMI - Pressure/Tension","IM_Percentage",
    "Motivation_Type","Campus_transformed","AT 3: Critical appraisal essay Wk9")

# Calculate the correlation matrix
correlation_matrix <- cor(regression_data[,numerical_vars])

# Create a correlation matrix plot using corrplot with labels
corrplot(correlation_matrix, method = "color", type = "upper", tl.col = "black", tl.srt = 45,
         tl.cex = 0.7, addCoef.col = "black", number.cex = 0.5)

```


```{r}
# Function to calculate RMSE for a linear regression model
calculate_RMSE <- function(model, data, target_column) {
  # Step 1: Predict values using the model
  predicted_values <- predict(model, newdata = data)
  
  # Step 2: Calculate the residuals
  residuals <- data[[target_column]] - predicted_values
  
  # Step 3: Square the residuals
  squared_residuals <- residuals^2
  
  # Step 4: Take the mean of squared residuals
  mean_squared_residuals <- mean(squared_residuals)
  
  # Step 5: Take the square root
  RMSE <- sqrt(mean_squared_residuals)
  
  # Print RMSE
  print(paste("Root Mean Square Error (RMSE): ", RMSE))
}

#function to get regression details
get_regression_summary <- function(model) {
  # Extract coefficients, standard errors, and p-values
  coefficients <- coef(summary(model))
  
  # Calculate the 95% confidence intervals
  CI_lower <- coefficients[, "Estimate"] - 1.96 * coefficients[, "Std. Error"]
  CI_upper <- coefficients[, "Estimate"] + 1.96 * coefficients[, "Std. Error"]
  
  # Combine the results into a data frame
  results <- data.frame(
    Coefficient = coefficients[, "Estimate"],
    CI_Lower = CI_lower,
    CI_Upper = CI_upper,
    P_Value = coefficients[, "Pr(>|t|)"]
  )
  
  return(results)
}

#function to calculate rsq for trees
calculate_rsq <- function(model, data, target_column) {
  # Predict using the model
  predicted_values <- predict(model, data)
  
  # Calculate TSS and RSS
  TSS <- sum((data[[target_column]] - mean(data[[target_column]]))^2)
  RSS <- sum((data[[target_column]] - predicted_values)^2)
  
  # Calculate R^2
  R2 <- (RSS/TSS)
  
  return(R2)
}
#function to calculate adj rsq for trees
calculate_adj_rsq<-function(data,k,model,target_column){
  R2<-calculate_rsq(model,data,target_column)
  adj.rsq<-1-((1-R2)*(nrow(data)/(nrow(data)-k-1)))
  return(adj.rsq)
}
```


```{r}
# Perform linear regression
model <- lm(`AT 3: Critical appraisal essay Wk9` ~ .-1,data = lin_regression_data)

# Print the model summary
summary(model)

get_regression_summary(model)
# Generate diagnostic plots
par(mfrow=c(2,2))  # Arrange plots in a 2x2 grid
plot(model)
calculate_RMSE(model, lin_regression_data, 'AT 3: Critical appraisal essay Wk9')
```


```{r}
# Perform stepwise regression
stepwise_model <- step(model, direction = "both")

# Summary of the selected model
summary(stepwise_model)

# Generate diagnostic plots
par(mfrow=c(2,2))  # Arrange plots in a 2x2 grid
plot(stepwise_model)
calculate_RMSE(stepwise_model, lin_regression_data, 'AT 3: Critical appraisal essay Wk9')
```


```{r}
# Fit the model with interaction terms
linear_interaction_model <- lm(`AT 3: Critical appraisal essay Wk9` ~ (Mode * Motivation_Type + ATAR*Motivation_Type+ATAR*`AT2 article summary Wk 5?`)-1, data = lin_regression_data)



# # Summary of the model
summary(linear_interaction_model)

get_regression_summary(linear_interaction_model)
# # Generate diagnostic plots
par(mfrow=c(2,2))  # Arrange plots in a 2x2 grid
plot(linear_interaction_model)

calculate_RMSE(linear_interaction_model, lin_regression_data, 'AT 3: Critical appraisal essay Wk9')

```


```{r}
# List of IMI subscales vs AT3 for moderator analysis
regression_data_mod <- dplyr::select(regression_data,-Age)

imi_subscales <- c("IMI - Interest/Enjoyment","IMI - Perceived Competence","IMI - Effort/Importance","IMI - Pressure/Tension","IMI - Perceived Choice","IMI - Value/Usefulness","IMI - Relatedness")

# Loop through each IMI subscale and run the moderator analysis
for(subscale in imi_subscales) {
  formula_str <- paste("`AT 3: Critical appraisal essay Wk9` ~ ATAR * `", subscale, "`","-1", sep = "")
  fit <- lm(formula_str, data = regression_data_mod)  # Note the data argument here
  # Call the calculate_RMSE function for each fit
  print(paste("Moderator Analysis for", subscale))
  print(summary(fit))
  print(get_regression_summary(fit))
  calculate_RMSE(fit, regression_data_mod, "AT 3: Critical appraisal essay Wk9")
}

```


```{r}
# Build the regression tree
tree_model <- rpart(`AT 3: Critical appraisal essay Wk9` ~ .,
                    data = lin_regression_data,
                    method = "anova")
# Plot the decision tree with minimum sample sizes
rpart.plot(tree_model, type = 1, extra = 101)
# Calculate and print R^2 using the custom function
R2_value <- calculate_rsq(tree_model, lin_regression_data, 'AT 3: Critical appraisal essay Wk9')
cat("R-squared:", R2_value, "\n")
Adj.RSQ<-calculate_adj_rsq(lin_regression_data,ncol(lin_regression_data) - 1,tree_model,'AT 3: Critical appraisal essay Wk9')
cat("Adj.R-squared:", Adj.RSQ, "\n")
calculate_RMSE(tree_model,lin_regression_data,'AT 3: Critical appraisal essay Wk9')
printcp(tree_model)
# Calculate and print adjusted R-squared
low_cp<-tree_model$cptable[which.min(tree_model$cptable[,"xerror"]),"CP"]
#This Model of Dtree is better than pruned tree bcoz it has lower RMSE and compared to lin regression also this is better Model.
```


```{r}
pruned_tree <- prune(tree_model, cp=0.078368)
rpart.plot(pruned_tree, type = 1, extra = 101)
R2_value <- calculate_rsq(pruned_tree, lin_regression_data, 'AT 3: Critical appraisal essay Wk9')
cat("R-squared:", R2_value, "\n")
Adj.RSQ<-calculate_adj_rsq(lin_regression_data,ncol(lin_regression_data) - 1,pruned_tree,'AT 3: Critical appraisal essay Wk9')
cat("Adj.R-squared:", Adj.RSQ, "\n")
calculate_RMSE(pruned_tree,lin_regression_data,'AT 3: Critical appraisal essay Wk9')
```

```{r}
#Select impotrant variables for linear regression using decision tree
variable_importance <- tree_model$variable.importance
print(variable_importance)



# Train linear regression model with important variables fron Decison tree Regression
best_lin_model <- lm(`AT 3: Critical appraisal essay Wk9` ~ (ATAR +IM_Percentage+`AT2 article summary Wk 5?`+`IMI - Pressure/Tension`+Campus_transformed)-1,data = lin_regression_data)


# Print the model summary
summary(best_lin_model)
calculate_RMSE(best_lin_model, lin_regression_data, 'AT 3: Critical appraisal essay Wk9')
get_regression_summary(best_lin_model)
# Generate diagnostic plots
par(mfrow=c(2,2))  # Arrange plots in a 2x2 grid
plot(best_lin_model)
```

#SVM regression
```{r}
model_svm_linear <- svm(`AT 3: Critical appraisal essay Wk9` ~ .,data = lin_regression_data, type = "eps-regression", kernel = "linear")
predictions <- predict(model_svm_linear, newdata =lin_regression_data)
summary(model_svm_linear)


# Assuming you have already fitted the SVM model and made predictions
# Calculate residuals
residuals <- predictions - lin_regression_data$`AT 3: Critical appraisal essay Wk9`

# Calculate MSE
mse <- mean(residuals^2)

# Calculate RMSE
rmse <- sqrt(mse)
print(rmse)
```


```{r}
# automated tuning of gamma paramter for RBF kernel
tuned_model <- tune.svm(`AT 3: Critical appraisal essay Wk9` ~ .,data = lin_regression_data, type = "eps-regression", kernel = "radial", gamma = 10^(-3:2))
best_model <- tuned_model$best.model
predictions <- predict(best_model, newdata =lin_regression_data)
```
```{r}
# Calculate residuals of svm
residuals <- predictions - lin_regression_data$`AT 3: Critical appraisal essay Wk9`

# Calculate MSE
mse <- mean(residuals^2)

# Calculate RMSE
rmse <- sqrt(mse)
print(rmse)
```

#Logistic Regression
```{r}
log_regression_data<-dplyr::select(regression_data,-c(Campus,`Region origin`,`AT 3: Critical appraisal essay Wk9`,AT3_Performance,TOTAL,TOTAL_Performance,SUS,`IMI - Interest/Enjoyment`,`IMI - Perceived Competence`,`IMI - Effort/Importance`,`IMI - Perceived Choice`,`IMI - Value/Usefulness`,`IMI - Relatedness`,`IMI Total`,EM_Percentage,Like_Hybrid,id,`Course Weighted Average`,Age))


model_log <- glm(AT3_Median_Performance ~.-1,data = log_regression_data,family = binomial)

# Print the model summary
summary(model_log)
```


```{r}
# Load necessary libraries


# Assuming you have a logistic regression model named 'model_log'
# Assuming 'regression_data' is your dataframe with the test data

# Make predictions using the logistic regression model
predictions <- predict(model_log, newdata = log_regression_data, type = "response")

# Convert probabilities to class predictions
predicted_classes <- ifelse(predictions > 0.5, 1,0)

# Create a data frame with actual and predicted class labels
confusion_data <- data.frame(Actual = regression_data$AT3_Median_Performance,
                             Predicted = predicted_classes)

print(paste("Accuracy of Basic Logistic Model is",(sum(confusion_data$Actual==confusion_data$Predicted)/nrow(confusion_data))*100))
```



```{r}

# Perform stepwise logistic regression
stepwise_log_model <- step(model_log, ,family = binomial)

# Summary of the selected model
summary(stepwise_log_model)
```

```{r}
# Make predictions using the logistic regression model
predictions <- predict(stepwise_log_model, newdata = log_regression_data, type = "response")

# Convert probabilities to class predictions
predicted_classes <- ifelse(predictions > 0.5, 1,0)

# Create a data frame with actual and predicted class labels
confusion_data <- data.frame(Actual = regression_data$AT3_Median_Performance,
                             Predicted = predicted_classes)

print(paste("Accuracy of Stepwise Logistic Model is",(sum(confusion_data$Actual==confusion_data$Predicted)/nrow(confusion_data))*100))
```



```{r}
# Assess variable importance using varImp
variable_importance_log <- varImp(model_log)
print(variable_importance_log)

```


```{r}
# Train logistic regression model with important variables
model_log <- glm(`AT3_Median_Performance` ~ (`ATAR`+Age_Group+Gender+`IMI - Pressure/Tension`+`AT2 article summary Wk 5?`+IM_Percentage+Motivation_Type+Campus_transformed)-1,data = log_regression_data,family = binomial)

# Print the model summary
summary(model_log)

```

```{r}
# Make predictions using the logistic regression model
predictions <- predict(model_log, newdata = log_regression_data, type = "response")

# Convert probabilities to class predictions
predicted_classes <- ifelse(predictions > 0.5, 1,0)

# Create a data frame with actual and predicted class labels
confusion_data <- data.frame(Actual = log_regression_data$AT3_Median_Performance,
                             Predicted = predicted_classes)

print(paste("Accuracy of Logistic Model with Important Variables is",(sum(confusion_data$Actual==confusion_data$Predicted)/nrow(confusion_data))*100))

```


```{r}
# Logistic Regression with Selected Two-Way Interaction Terms
interaction_model <- glm(AT3_Median_Performance ~ (ATAR + `AT2 article summary Wk 5?` + IM_Percentage + ATAR:`AT2 article summary Wk 5?` + ATAR:IM_Percentage + `AT2 article summary Wk 5?`:IM_Percentage)-1, data = log_regression_data, family = binomial)

# Summary of the model
summary(interaction_model)

```


```{r}
# Make predictions using the logistic regression model
predictions <- predict(interaction_model, newdata = log_regression_data, type = "response")

# Convert probabilities to class predictions
predicted_classes <- ifelse(predictions > 0.5, 1,0)

# Create a data frame with actual and predicted class labels
confusion_data <- data.frame(Actual = log_regression_data$AT3_Median_Performance,
                             Predicted = predicted_classes)

print(paste("Accuracy of Logistic Regression using interactions",(sum(confusion_data$Actual==confusion_data$Predicted)/nrow(confusion_data))*100))
```

```{r}
# Logistic Regression with Only Significant Interaction Term
significant_logistic_interactions <- glm(AT3_Median_Performance ~ (ATAR + `AT2 article summary Wk 5?` + ATAR:`AT2 article summary Wk 5?`)-1, 
                                         data = log_regression_data, 
                                         family = binomial)

# Summary of the model
summary(significant_logistic_interactions)

```


```{r}
# Make predictions using the logistic regression model
predictions <- predict(significant_logistic_interactions, newdata = log_regression_data, type = "response")

# Convert probabilities to class predictions
predicted_classes <- ifelse(predictions > 0.5, 1,0)

# Create a data frame with actual and predicted class labels
confusion_data <- data.frame(Actual = log_regression_data$AT3_Median_Performance,
                             Predicted = predicted_classes)

print(paste("Accuracy of Logistic Regression using significant interactions",(sum(confusion_data$Actual==confusion_data$Predicted)/nrow(confusion_data))*100))
```


#SVM classification
```{r}
model_svm_class <- svm(AT3_Median_Performance ~.,data = log_regression_data,type = "C-classification",kernel = "linear")
predictions <- predict(model_svm_class, newdata =log_regression_data)
# Evaluate the model
# Calculate accuracy
accuracy <- sum(predictions == log_regression_data$AT3_Median_Performance) / length(predictions)
print(accuracy*100)
```

```{r}
# automated tuning of gamma paramter for RBF kernel
log_regression_data$AT3_Median_Performance <- as.factor(log_regression_data$AT3_Median_Performance)
tuned_model <- tune.svm(AT3_Median_Performance ~.,data = log_regression_data,type = "C-classification", kernel = "radial", gamma = 10^(-3:2))
best_model <- tuned_model$best.model
predictions <- predict(best_model, newdata =log_regression_data)
# Evaluate the model
# Calculate accuracy
accuracy <- sum(predictions == log_regression_data$AT3_Median_Performance) / length(predictions)
print(accuracy*100)
```

```{r}
write_xlsx(lin_regression_data, path = "C:\\Users\\Angat\\Downloads\\hybrid_teaching_cleaned_transformed.xlsx")
```